---
title: "Assignment 3"
author: "Insert My Name"
date: "2023-10-09"
output: html_document # you can change to other output format if you want
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Exploratory data analysis
```{r}
library(Stat2Data)
data("Hawks")
head(Hawks)
```


## 1.1 (Q1)   

<!-- Include your answers for each question under a second level heading eg ## 1.1 (Q1) -->
**** insert your solutions here *****
```{r}
HawksTail <- Hawks[["Tail"]]
head(HawksTail)
print(mean(HawksTail))
print(median(HawksTail))

```


## 1.2 (Q1)
```{r}
library(dplyr)
Hawksum <- Hawks %>%  
  summarize(
  Wing_mean = mean(Wing,na.rm = TRUE),
  Wing_t_mean = mean(Wing,t=0.5,na.rm = TRUE),
  Wing_med = median(Wing,na.rm = TRUE),
  Weight_mean = mean(Weight,na.rm = TRUE),
  Weight_t_mean = mean(Weight,t=0.5,na.rm = TRUE),
  Weight_med = median(Weight,na.rm = TRUE))

head(Hawksum)
```

## 1.2 (Q2)
```{r}
Hawksum1 <- Hawks %>%
group_by(Species) %>%
summarise(
  Wing_mean = mean(Wing,na.rm = TRUE),
  Wing_t_mean = mean(Wing,t=0.5,na.rm = TRUE),
  Wing_med = median(Wing,na.rm = TRUE),
  Weight_mean = mean(Weight,na.rm = TRUE),
  Weight_t_mean = mean(Weight,t=0.5,na.rm = TRUE),
  Weight_med = median(Weight,na.rm = TRUE))

head(Hawksum1)
```


## 1.3 (Q1)
aA+b
```{r}
print(2*mean(HawksTail) + 3)
print(mean(2*HawksTail+3))
```


## 1.3 (Q2)
a^2p
|a|q
```{r}
print(4*var(HawksTail))
print(var(2*HawksTail+3))

print(2*sd(HawksTail))
print(sd(2*HawksTail+3))
```
## 1.4 (Q1)
```{r}
hal<-Hawks$Hallux # Extract the vector of hallux lengths
hal<-hal[!is.na(hal)] # Remove any nans

outlier_val<-100
num_outliers<-10
corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))

mean(hal)
mean(corrupted_hal)


num_outliers_vect <- seq(0,1000)
means_vect <- c()
for(num_outliers in num_outliers_vect){
corrupted_hal <- c(hal,rep(outlier_val,times=num_outliers))
means_vect <- c(means_vect, mean(corrupted_hal))
}

head(means_vect)

```
```{r}
num_outliers_vect <- seq(0,1000)
medians_vect <-c()
for(num_outliers in num_outliers_vect){
corrupted_hal <- c(hal,rep(outlier_val,times=num_outliers))
medians_vect <-c(medians_vect, median(corrupted_hal))
}
head(medians_vect)
```



## 1.4 (Q2)
```{r}
num_outliers_vect <- seq(0,1000)
t_means_vect <-c()
for(num_outliers in num_outliers_vect){
corrupted_hal <- c(hal,rep(outlier_val,times=num_outliers))
t_means_vect <- c(t_means_vect, mean(corrupted_hal,t=0.1))
}

head(t_means_vect)
```

## 1.4 (Q3)
```{r}
library(tidyr)
library(ggplot2)



df_means_medians <- data.frame(num_outliers=num_outliers_vect, mean=means_vect,
t_mean=t_means_vect, median=medians_vect)

df_means_medians %>%
pivot_longer(!num_outliers, names_to = "Estimator", values_to = "Value") %>%
ggplot(aes(x=num_outliers,color=Estimator, linetype=Estimator,y=Value)) +
geom_line()+xlab("Number of outliers")
```


## 1.5 (Q1)
```{r}
ggplot(data=Hawks,aes(x=Species, y=Weight)) +
geom_boxplot()+xlab("Species") + ylab("Weight")
```

## 1.5 (Q2)
```{r}
probabilities <- c(0.25, 0.5, 0.75)

quantiles <- Hawks %>% 
  select(Weight, Species) %>%
  filter(!is.na(Weight) & is.numeric(Weight)) %>%
  group_by(Species) %>%
  summarize(
    quantile025 = quantile(Weight, probs = probabilities[1], na.rm = TRUE),
    quantile050 = quantile(Weight, probs = probabilities[2], na.rm = TRUE),
    quantile075 = quantile(Weight, probs = probabilities[3], na.rm = TRUE)
  )


quantiles 




```

## 1.5 (Q3)
```{r}
num_outliers <- function(x) {
  q25 <- quantile(x, probs = 0.25, na.rm = TRUE)
  q75 <- quantile(x, probs = 0.75, na.rm = TRUE)
  iqr <- q75 - q25
  outliers <- sum(x < q25 - 1.5 * iqr | x > q75 + 1.5 * iqr, na.rm = TRUE)
  return(outliers)
}

# Test the function
num_outliers(c(0, 40, 60, 185))


```
## 1.5 (Q4)
```{r}
num_outliers_table <- Hawks %>%
  group_by(Species) %>%
  summarise(
    num_outliers_weight = num_outliers(Weight)
  )

num_outliers_table

```

## 1.6 (Q1)
```{r}
Covar <- cov(Hawks$Weight, Hawks$Wing, use = "complete.obs")
Corre <- cor(Hawks$Weight, Hawks$Wing, use = "complete.obs")

print(Covar)
print(Corre)
```

## 1.6 (Q2)
S' = acS
R' = S/sd(x)*sd(y)
```{r}
a <- 2.4
b <- 7.1
c <- -1
d <- 3
X <- Hawks$Weight
Y <- Hawks$Wing
X_tilde <- a * Hawks$Weight + b
Y_tilde <- c * Hawks$Wing + d

print(a*c*cov(X, Y,use = "complete.obs"))
print(cov(X_tilde, Y_tilde,use = "complete.obs"))

print(cor(X, Y,use = "complete.obs"))
print(cor(X_tilde, Y_tilde,use = "complete.obs"))


```






# 2. Random experiments, events and sample spaces, and the set theory
## 2.1 (Q1)
The concepts of random experiments, event, and sample space
1. A random experiment is a procedure (real or imagined) which:
a). has a well-defined set of possible outcomes;
b). could (at least in principle) be repeated arbitrarily many times.
Examples: A coin flip for a coin (outcomes: Heads up, tails up)
2. An event is a set (i.e. a collection) of possible outcomes of an experiment
Example: heads up; tails up; heads up or tails up
It is important to note that an event is different from a outcome (although an event may
contain a single outcome).
3. A sample space is the set of all possible outcomes of interest for a random experiment
Example: heads up, tails up



## 2.1 (Q2)
Event: Getting a sum of 7 on the two rolls. This event is represented as {(1, 6), (2, 5), (3, 4), (4, 3), (5, 2), (6, 1)}.

The sample space for this random experiment can be written as a set Ω, which contains all possible outcomes. The sample space Ω in this case consists of all possible pairs of outcomes from two rolls of a six-sided die:

Ω = {(1, 1), (1, 2), (1, 3), (1, 4), (1, 5), (1, 6), (2, 1), (2, 2), ..., (6, 5), (6, 6)}, the total number is 36

The empty set (∅), which represents an event with no outcomes, can be considered as an event. It's referred to as the "null event" or the "impossible event." 

## 2.2 (Q1)
AUB = {1,2,3,4,6}; AUC ={1,2,3,4,5,6}
A∩B = {2} A∩C = {}
A\B = {1,3} A\C ={1,2,4}
No, Yes
Yes
{1,2,4},{3,5,6}; {1,2},{3,4},{5,6}
## 2.2 (Q2)
Expression for (A^c)^c:
The complement of a complement is the original set itself. So, (A^c)^c = A.

What is Ω^c:
The complement of the entire sample space Ω is the empty set (the set with no elements), denoted as ∅. So, Ω^c = ∅.

Subsets:
If A ⊆ B (A is a subset of B), then for any element x in the complement of B (x ∈ B^c), it must be the case that x is also in the complement of A (x ∈ A^c). In other words, B^c ⊆ A^c.

De Morgan's Laws:
a. Intersection Law: (A ∩ B)^c = A^c ∪ B^c. This means that the complement of the intersection of two events is equal to the union of their complements.
b. General Case for a Sequence of Events: For a sequence of events A1, A2, ..., AK ⊆ Ω, the complement of the intersection of these events is the union of the complements: (∩(from k=1 to K) Ak)^c = ∪(from k=1 to K) A^c.

De Morgan's Laws:
a. Union Law: (A ∪ B)^c = A^c ∩ B^c. This means that the complement of the union of two events is equal to the intersection of their complements.
b. General Case for a Sequence of Events: For a sequence of events A1, A2, ..., AK ⊆ Ω, the complement of the union of these events is the intersection of the complements: (∪(from k=1 to K) Ak)^c = ∩(from k=1 to K) A^c.
## 2.2 (Q3)
2^k
## 2.2 (Q4)
 A = ∅
 
The events S1, S2, S3, S4 form a partition of A1 ∪ A2 ∪ A3 ∪ A4 if:
They are mutually exclusive (disjoint): For all i ≠ j, Si ∩ Sj = ∅, which means that no two of them share any elements.
Their union covers the entire set A1 ∪ A2 ∪ A3 ∪ A4: ∪(i=1 to 4) Si = A1 ∪ A2 ∪ A3 ∪ A4. This means that the union of the S events includes all elements from the A events.
## 2.2 (Q5)
1A^c = 1- 1A
B = A ∪ A^c


## 2.2 (Q6)

# 3. Visualisation

## 3 (Q1)
```{r}
ggplot(data=Hawks, aes(x=Tail, color=Species))+
  geom_density()+theme_bw()+xlab("Tail (mm)")+ylab("Density")
```

## 3 (Q2)
```{r}
library(ggplot2)

ggplot(data = Hawks, aes(x = Tail, y = Species, color = Species, fill = Species)) +
  geom_violin(color = "black", size = 0.5) +  # 设置边框为黑色
  theme_bw() +
  xlab("Tail (mm)") +
  ylab("Density")

```

## 3 (Q3)
```{r}
ggplot(data = Hawks, aes(y = Weight, x = Tail, color = Species, shape = Species)) +
  geom_point(size = 2) +
  xlab("Tail (mm)") +
  ylab("Weight (mm)")
  
```

## 3 (Q4)
```{r}

ggplot(data = Hawks, aes(y = Weight, x = Tail, color = Species)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~Species, scales = "free") +
  xlab("Tail (mm)") +
  ylab("Weight (mm)") +
  theme_minimal()


```

## 3 (Q5)

# Find the heaviest hawk
heaviest_hawk <- Hawks %>%
  filter(Weight == max(Weight, na.rm = TRUE)) %>%
  select(Weight, Tail, Species)

# Your existing plot code
ggplot(data = Hawks, aes(y = Weight, x = Tail, color = Species, shape = Species)) +
  geom_point(size = 2) +
  xlab("Tail (mm)") +
  ylab("Weight (mm)") +

```{r}
library(ggplot2)


# Find the heaviest hawk
heaviest_hawk <- Hawks %>%
  filter(Weight == max(Weight, na.rm = TRUE)) %>%
  select(Weight, Tail, Species)

# Create the scatter plot
plot <- ggplot(data = Hawks, aes(y = Weight, x = Tail, color = Species)) +
  geom_point(size = 2) +
  xlab("Tail (mm)") +
  ylab("Weight (mm)")

plot_with_arrow <- plot +
  geom_curve(data = heaviest_hawk, aes(x = Tail, y = Weight - 200, xend = Tail, yend = Weight),
             arrow = arrow(type = "closed", length = unit(0.1, "inches"), ends = "last", angle = 15),
             color = "black") +
  geom_text(data = heaviest_hawk, aes(label = "heaviest hawk", x = Tail, y = Weight - 220),color ="black") +
  theme(legend.position = "none")

# Print the plot with the arrow and annotation
print(plot_with_arrow)



```

